01 - TreehopperFitness - Quality control and data prep

### 2023-08-25 scripts/Publilia-01-qc.Rmd
init `Publilia-04-within_clutch_iteroparity.Rmd`

## 10.1 - Generate data for fecundity modeling

```{r}
fec_model_data <- visited_1st_4d_data_tb %>%
  filter(!is.na(guard_duration)) %>%
  group_by(location) %>%
  mutate(eggs_ever_visited = any(!mom_at_init & visit_ends <= est_hatch)) %>%
  ungroup() %>%
  select(Female_ID, location, guard_duration, est_rel2h, initiation_date, est_hatch,
         mom_at_init, clutch_initiated_order, initL, initS,
         visit_starts, visit_ends, last_day_fID,
         num_at_init, n_clutches_on_plant, `visited in 4 days`, eggs_ever_visited, ants,
         females_removed, ants_removed, 
         max_egg_num, egg_fate, egg_fate_date)
```

## 10.2 - Calculate median of top 5 egg counts in lieu of max egg counts
Taking the median (or mean) of the top 5 egg_num counts yields an estimate of max egg num that is more robust to outliers/miscounts
```{r, warning=FALSE}
tmp_t5eggs_data <- clutch_status_init_hatch_labels_v1 %>%
  mutate(est_hatch = as_date(ifelse(is.na(hatch_date),
                                 initiation_date + 20,
                                 hatch_date)),
         clutch_age = as.numeric(Date-initiation_date)) %>%
  # remove duplicate egg counts due to multiple females at a location on a day
  distinct(location, Date, initiation_date, est_hatch, clutch_age,
           egg_num, egg_len) %>%
  group_by(location) %>%
    arrange(desc(egg_num)) %>%
    mutate(top5egg_counts = rank(desc(egg_num), ties.method = "first") <= 5,
           top5egg_lens = rank(desc(egg_len), ties.method = "first") <= 5) %>%
    #filter(location == "A13_11") %>%
    select(location, Date, initiation_date, est_hatch, clutch_age,
           egg_num, egg_len, top5egg_counts, top5egg_lens) %>%
    mutate(med_t5eggs = median(ifelse(top5egg_counts,egg_num,NA), na.rm=T),
           mean_t5eggs = mean(ifelse(top5egg_counts,egg_num,NA), na.rm=T),
           max_n_eggs = max(ifelse(top5egg_counts,egg_num,NA), na.rm=T),
           med_l_t5eggs = median(ifelse(top5egg_lens,egg_len,NA), na.rm=T),
           mean_l_t5eggs = mean(ifelse(top5egg_lens,egg_len,NA), na.rm=T),
           max_l_eggs = max(ifelse(top5egg_lens,egg_len,NA), na.rm=T)) %>%
  ungroup()

top5eggs_summary <- tmp_t5eggs_data %>%
  group_by(location) %>%
    summarize(med_t5eggs = first(med_t5eggs),
              mean_t5eggs = first(mean_t5eggs),
              max_n_eggs = first(max_n_eggs),
              med_l_t5eggs = first(med_l_t5eggs),
              mean_l_t5eggs = first(mean_l_t5eggs),
              max_l_eggs = first(max_l_eggs)) %>%
  ungroup()
```

@@

>>>>

## 10.1 - Calculate median of top 5 egg counts in lieu of max egg counts
Taking the median (or mean) of the top 5 egg_num counts yields an estimate of max egg num that is more robust to outliers/miscounts
```{r, warning=FALSE}
t5eggs_data <- clutch_status_init_hatch_labels_v1 %>%
  mutate(est_hatch = as_date(ifelse(is.na(hatch_date),
                                 initiation_date + 20,
                                 hatch_date)),
         clutch_age = as.numeric(Date-initiation_date)) %>%
  # remove duplicate egg counts due to multiple females at a location on a day
  distinct(location, Date, initiation_date, est_hatch, clutch_age,
           egg_num, egg_len) %>%
  group_by(location) %>%
    arrange(desc(egg_num)) %>%
    mutate(top5egg_counts = rank(desc(egg_num), ties.method = "first") <= 5,
           top5egg_lens = rank(desc(egg_len), ties.method = "first") <= 5) %>%
    #filter(location == "A13_11") %>%
    select(location, Date, initiation_date, est_hatch, clutch_age,
           egg_num, egg_len, top5egg_counts, top5egg_lens) %>%
    mutate(med_t5eggs = median(ifelse(top5egg_counts,egg_num,NA), na.rm=T),
           mean_t5eggs = mean(ifelse(top5egg_counts,egg_num,NA), na.rm=T),
           max_n_eggs = max(ifelse(top5egg_counts,egg_num,NA), na.rm=T),
           med_l_t5eggs = median(ifelse(top5egg_lens,egg_len,NA), na.rm=T),
           mean_l_t5eggs = mean(ifelse(top5egg_lens,egg_len,NA), na.rm=T),
           max_l_eggs = max(ifelse(top5egg_lens,egg_len,NA), na.rm=T)) %>%
  ungroup()

top5eggs_summary <- t5eggs_data %>%
  group_by(location) %>%
    summarize(med_t5eggs = first(med_t5eggs),
              mean_t5eggs = first(mean_t5eggs),
              max_n_eggs = first(max_n_eggs),
              med_l_t5eggs = first(med_l_t5eggs),
              mean_l_t5eggs = first(mean_l_t5eggs),
              max_l_eggs = first(max_l_eggs)) %>%
  ungroup()
```

## 10.2 - Generate data for fecundity modeling

```{r}
fec_model_data <- visited_1st_4d_data_tb %>%
  filter(!is.na(guard_duration)) %>%
  group_by(location) %>%
  mutate(eggs_ever_visited = any(!mom_at_init & visit_ends <= est_hatch)) %>%
  ungroup() %>%
  select(Female_ID, location, guard_duration, est_rel2h, initiation_date, est_hatch,
         mom_at_init, clutch_initiated_order, initL, initS,
         visit_starts, visit_ends, last_day_fID,
         num_at_init, n_clutches_on_plant, `visited in 4 days`, eggs_ever_visited, ants,
         females_removed, ants_removed, 
         max_egg_num, egg_fate, egg_fate_date) %>%
  left_join(top5eggs_summary)
```

@@

# 13 - Write data for `Publilia-04-within_clutch_iteroparity.Rmd` to `.rds`

## `Publilia-04-within_clutch_iteroparity.Rmd`
```{r}
write_rds(clutch_egg_num_data, here("saved_rds/clutch_egg_num_data"))
write_rds(t5eggs_data, here("saved_rds/t5eggs_data"))
write_rds(fec_model_data, here("saved_rds/fec_model_data.rds"))
write_rds(clutch_status_init_hatch_labels_v1, here("saved_rds/clutch_status_init_hatch_labels_v1.rds"))
```
